### 2022-09-16 paper/abstract.tex
Paper notes

Many data science problems require processing log data derived from web pages, apis or other 
internet based services. URLs are one of the key pieces of information that describe almost
all types of internet activity. While they are structurally rich, it is not entirely obvious 
how to derive the optimal features from a set of URLs or domains.

In this talk we will discuss a wide variety of approaches to deriving features from URLs and domains. 
We begin with a breakdown of the broad categories of URL derived features and then discuss some of 
the key implementation details, and types of problems for which they are appropriate.
Many of the ideas in this talk have been implemented in the open source python library url2features.

https://github.com/getting-data-science-done/url2features